An upper bound to gas storage and delivery via pressure-swing adsorption in porous materials

### 2020-06-01 gas-adsorption/gas-ads-appendix.tex
rework appendix D to hopefully fit better with new figure

\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
Every MOF has a $\Delta g_\text{st}$ at $\pfull$ and $\pempty$ corresponding
to a full and empty density. The deliverable capacity of the MOF is equal to
the difference between the full and empty density. Examining
Fig.~\ref{fig:methane-gst}, we see that a significant variety of known
experimental $\Delta g_\text{st}$ curves are monotonic. Furthermore, our
qualitative argument in Section~\ref{sec:upper-bound} suggests that this
function \emph{should} monotonically increase for rigid substrates, as
increasing the density of gas causes some of the gas to reside in higher-energy
sites. If this is always the case, the deliverable capacity of a real material (with a non-spatially uniform potential) with a given $\gst$ at $\pfull$ or $\pempty$ is bounded below by the deliverable capacity in our idealized substrate with $\V=\gst$, even for a value of $\V$ that is not optimal

To show this, we perform a thought experiment illustrated in
Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of volumes containing
gas. Within each pair, one volume is filled with a porous material, while the
other offers a uniform potential energy $\V$ (our idealized substrate). In the
first (second) pair, the density of gas in each volume is $\rhofull$
($\rhoempty$), corresponding to the density of gas adsorbed in the porous
material at $\pfull$ ($\pempty$).

We now consider what happens if we introduce a diffusive connection between a
volume of MOF and a volume with an idealized substrate that initially contain
the same density of gas, for instance by connecting them with a tube. If the
chemical potential of the two volumes are equal, their density will be
unchanged; otherwise gas will flow into the volume with lower chemical
potential. The difference in chemical potential is equal to the difference

\begin{align}
   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
   \\
   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
\end{align}
where $\mu_{\text{MOF}}(\rho)$ and $\mu_{\V}(\rho)$ are the chemical potentials
of gas in the porous material and ideal substrate with potential $\V$, respectively. Thus if $\gst(\rhoempty) =
\V$ then the low-density containers will remain at their initial density after
they are connected. Thus, the deliverable capacity of the MOF will be greater
than the deliverable capacity of the ideal substrate with uniform potential $\V$ if and
only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$ does \emph{not} monotonically
increase, since the gas in the high-density system will flow from the ideal
substrate to the MOF. By the same token, if we consider the case where
$\gst(\rhofull) = \V$, then for the MOF to achieve a greater deliverable
capacity than the idealized substrate, the MOF must have less residual gas,
which means that gas must spontaneously flow from the low-density MOF to the
volume with potential $\V$, which means that $\gst(\rhoempty)>\V$. Once again,
exceeding our upper bound requires a material with a $\gst(\rho)$ that does not
increase monotonically.

>>>>

\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
Every MOF has a $\Delta g_\text{st}$ at $\pfull$ and $\pempty$ corresponding
to a full and empty density. The deliverable capacity of the MOF is equal to
the difference between the full and empty density. Examining
Fig.~\ref{fig:methane-gst}, we see that a significant variety of known
experimental $\Delta g_\text{st}$ curves are monotonic. Furthermore, our
qualitative argument in Section~\ref{sec:upper-bound} suggests that this
function \emph{should} monotonically increase for rigid substrates, as
increasing the density of gas causes some of the gas to reside in higher-energy
sites. If this is always the case, the deliverable capacity of a real material
(with a non-spatially uniform potential) with a given $\gst$ at $\pfull$ or
$\pempty$ is bounded below by the deliverable capacity in our idealized
substrate with $\V=\gst$, even for a value of $\V$ that is not optimal.

To show this, we perform a thought experiment illustrated in
Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of volumes.  Two of
these volumes contain a real adsorbent material.  These adsorbent volumes are
connected to bulk gas reservoirs at $\pfull$ and $\pempty$ respectively, and
therefore have densities $\rhofull$ and $\rhoempty$.  The other two volumes
volumes have equal density to the two adsorbent volumes ($\rhofull$ and $\rhoempty$), but contain our
idealized substrate with uniform potential energy $\V$.

We now consider what happens if we open a diffusive connection between a
volume of MOF and a volume with an idealized substrate that initially contain
the same density of gas, for instance by connecting them with a tube. If the
chemical potential in the idealized substrate is higher than in the porous material,
then gas will flow out of the idealized substrate, lowering its density.  Conversely
if the chemical potential is lower in the idealized substrate than in the
volume of porous material, then gas will flow into the idealized substrate, increasing
its density.
The difference in chemical potential between those two volumes is
\begin{align}
   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
   \\
   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
\end{align}
where $\mu_{\text{MOF}}(\rho)$ and $\mu_{\V}(\rho)$ are the chemical potentials
of gas in the porous material and ideal substrate with potential $\V$, respectively.
Thus if $\gst(\rhoempty) =
\V$ then the low-density containers will remain at their initial density after
they are connected. Thus, the deliverable capacity of the MOF will be greater
than the deliverable capacity of the ideal substrate with uniform potential $\V$ if and
only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$ does \emph{not} monotonically
increase, since the gas in the high-density system will flow from the ideal
substrate to the MOF. By the same token, if we consider the case where
$\gst(\rhofull) = \V$, then for the MOF to achieve a greater deliverable
capacity than the idealized substrate, the MOF must have less residual gas,
which means that gas must spontaneously flow from the low-density MOF to the
volume with potential $\V$, which means that $\gst(\rhoempty)>\V$. Once again,
exceeding our upper bound requires a material with a $\gst(\rho)$ that does not
increase monotonically.
